### 2024-01-23 notes_to_selves.tex
split

\documentclass{article}

\usepackage{soul}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsmath}

\begin{document}

\section{TODOs}

TODO:
   \begin{itemize}

      \item split the problems in two sets one for refining parameters and bounds, the other for overall evaluation
      \item read mangling paper in detail
      \item set up tables
      \item finish implementation (mostly there)
      \item test and evaluate implementation 
      \item (can wait) merge implementation (review code with active contributor(s) beforehand)
      \item find pretty inline setup for pseudo code
      \item (in order) write high/low level algo
      \item write implementation and evaluation
      \item write background
      \item write related work
      \item write conclusion
      \item write intro
      \item reread, spellcheck ...
    \end{itemize}

\begin{center}
\captionof{table}{Bounds and options to test}
\begin{tabular}{cccccc}
   \toprule
   \multicolumn{5}{c}{monomorphisation point (steps)} \\
   0 & 25 & 50 & 100 & (200) \\
   \midrule
   \multicolumn{4}{c}{new clauses multiplier} \\
   1.0 & 1.5 & 2.0 & (3.0) \\
   \midrule
   \multicolumn{3}{c}{total new clauses cap} \\
   2000 & 5000 & none\\
   \midrule
   \multicolumn{4}{c}{loop number} \\
   2 & 3 & 4 & 5 \\
   \midrule
   \multicolumn{5}{c}{monomorphisation point (steps)} \\
   0 & 25 & 50 & 100 & (200) \\
   \midrule
   \multicolumn{5}{c}{e-timeout} \\
   2.0 & 5.0 & 10.0 & 15.0 & no timeout \\
   \midrule
   \multicolumn{3}{c}{new poly ty arg limit per fun sym multiplier} \\
   0 & 0.5 & 1.0  \\
   \midrule
   \multicolumn{3}{c}{new poly ty arg limit per fun sym floor} \\
   0 & 2 & 5  \\
   \midrule
   \multicolumn{3}{c}{new poly ty arg limit per fun sym cap} \\
   10 & 20 & 50  \\
   \midrule
   \multicolumn{5}{c}{new mono ty arg limit per fun sym multiplier} \\
   0..5 & 1.0 & 1.5 & 2.0 & (3.0)\\
   \midrule
   \multicolumn{5}{c}{new mono ty arg limit per fun sym floor} \\
   0 & 2 & 5 & 7 & 10\\
   \midrule
   \multicolumn{4}{c}{new mono ty arg limit per fun sym cap} \\
   10 & 20 & 50 & 100  \\
   \midrule
   \multicolumn{3}{c}{monomorphising subst per clause} \\
   2 & 5 & 7  \\
   \midrule
   \multicolumn{5}{c}{subst per clause (total, split 50/50)} \\
   1000 & 2000 & 5000 & 10000 & 15000  \\

   \bottomrule
\end{tabular}
\end{center}

\end{document}