### 2024-02-09 notes_to_selves.tex
re-wrote a fair bit of the introduction

\documentclass{article}

\usepackage{soul}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsmath}

\begin{document}

\section{TODOs}

TODO:
   \begin{itemize}

      \item set up tables
      \item finish implementation (mostly there)
      \item test substitution sorting methods
      \item tune options
      \item compare with no mono zipperposition
      \item compare other provers
      \item (can wait) merge implementation (review code with active contributor(s) beforehand)
      \item find pretty inline setup for pseudo code
      \item reread, spellcheck ...
    \end{itemize}

\begin{center}
\captionof{table}{Bounds and options to test}
\begin{tabular}{cccccc}
   \toprule
   \multicolumn{5}{c}{monomorphisation point (steps)} \\
   0 & 25 & 50 & 100 & (200) \\
   \midrule
   \multicolumn{4}{c}{new clauses multiplier} \\
   1.0 & 1.5 & 2.0 & (3.0) \\
   \midrule
   \multicolumn{3}{c}{total new clauses cap} \\
   2000 & 5000 & none\\
   \midrule
   \multicolumn{4}{c}{loop number} \\
   2 & 3 & 4 & 5 \\
   \midrule
   \multicolumn{5}{c}{e-timeout} \\
   2.0 & 5.0 & 10.0 & 15.0 & no timeout \\
   \midrule
   \multicolumn{3}{c}{new poly ty arg limit per fun sym multiplier} \\
   0 & 0.5 & 1.0  \\
   \midrule
   \multicolumn{3}{c}{new poly ty arg limit per fun sym floor} \\
   0 & 2 & 5  \\
   \midrule
   \multicolumn{3}{c}{new poly ty arg limit per fun sym cap} \\
   10 & 20 & 50  \\
   \midrule
   \multicolumn{5}{c}{new mono ty arg limit per fun sym multiplier} \\
   0..5 & 1.0 & 1.5 & 2.0 & (3.0)\\
   \midrule
   \multicolumn{5}{c}{new mono ty arg limit per fun sym floor} \\
   0 & 2 & 5 & 7 & 10\\
   \midrule
   \multicolumn{4}{c}{new mono ty arg limit per fun sym cap} \\
   10 & 20 & 50 & 100  \\
   \midrule
   \multicolumn{3}{c}{monomorphising subst per clause} \\
   2 & 5 & 7  \\
   \midrule
   \multicolumn{5}{c}{subst per clause (total, split 50/50)} \\
   1000 & 2000 & 5000 & 10000 & 15000  \\
   \midrule
   \multicolumn{3}{c}{zipp mode} \\
   best & ho-pragmatic & ho-competitive \\

   \bottomrule
\end{tabular}
\end{center}

\end{document}