### 2025-04-16 REBUTTAL.txt
more

Dear Jasmin,

Thank you for your submission to CADE-30. The CADE-30 rebuttal period
has now started. The reviews on your paper are attached to this e-mail.
You have the opportunity to respond to them until April 18, midnight
(AoE). To submit your response please log on the EasyChair Web page
for CADE-30 and select your submission on the menu.

Please keep in mind the following during this process:

* We are not enforcing any upper bound on the length of your response
 but please try to limit it to 1000 words. Anything over that will
 be optional for the reviewers to read.

* The response must focus on factual errors in the reviews and the
 questions posed by the reviewers, if any. Please do not provide new
 research or reformulate the presentation. Try to be as concise and
 to the point as possible.

* The response period is an opportunity to react to the reviews but
 there is no requirement to do so. Thus, if you feel the reviews are
 accurate and the reviewers have not asked any questions, you do not
 have to respond. On the other hand, reviewers may not look favorably
 on a paper if authors fail to answer their questions.

* The reviews are as submitted by the PC members. Please keep in mind
 that this is not their final version and no coordination among the
 PC members was required at this stage. The reviews may later be
 updated to take into account your response and the discussions among
 the PC members. Moreover, we may find it necessary to solicit other
 outside reviews after the rebuttal period.

* While the program committee will read your response carefully and
 take it into account during the discussions, it will not directly
 respond to your response, either through the PC chairs or in the
 final versions of the reviews.

* Your response will be seen by all PC members who have access to the
 discussion of your paper, so please be polite and constructive.

Best wishes,
 Clark and Uwe

----------------------- REVIEW 1 ---------------------
SUBMISSION: 17
TITLE: Iterative Monomorphisation
AUTHORS: Tanguy Bozec and Jasmin Blanchette

----------- Overall evaluation -----------
The paper discusses the process of monomorphisation, that is
heuristic instantiation of a polymorphic automated reasoning
problem and then using a monomorphic prover on it. The authors
re-implement the procedure in Zipperposition and evaluate it.

I have two main concerns about the paper:

First, the novelty. The authors admit that algorithms
doing the same task in a similar way already existed in HOL
Light since the ninties and later in Sledgehammer. A more
detailed comparison of these would be needed. Apart from the
already mentioned monomorhpisation is also needed in many
other contexts, for example interation of ITPs with Metis and
leancop or in HOLyHammer in
HOL4. I do not know if some of these algorithms are the same,
but in any case I would really need to see a comparison of
the current work with the existing ones. What do other tools
that call monomorhpic ATPs like Why3 do? Even if some tools
might fail it should be discussed.

Second, it is not clear what logic does the work support.
Quote: "any reasonable rank-1 polymorphic logic", what is
reasonable? In the second definition the authors define a
"type arity" of a symbol and actually it is based on this
definition that I infer some things that are possible.
But does the logic have subtyping or even just type classes
(as Isabelle/HOL's does)? Things should be defined more
precisely.

Minor:
Page 1 'there exists, for any ... ' can we do it in the normal
order as in 'for any ... there exists ...'

'appears to be implemented' please look at what is implemented
and describe what things do what in HOL Light, HOL4, and other
logic transformation tools, say what does why3 or leo's
transformation preprocessor do with polymorphic problems.

Page 2: 'instances can be generated is infinite' it would be
good to show an example problem where more iterations are
needed to solve.

Page 3: 'applying a fixed number of iterations' why fixed?

Fig 2: The algorithm in MESON looks much simpler in code. Can you
compare them?

Page 7: 'we sketch an example' this is not really an example.

Why are the complexities useful? Are the complexities of the
Sledgehammer / MESON / ... algorithms different?

Fig 4: This does not bring much, the actual code would look
shorter than this algorithmization... Could we write these
algorithms more compactly?

Results (mostly table 1 on page 11):
Just an idea, as monomorphisation is supposed to add most to the
already existing polymorphic procedure, maybe rather than optimize
it on its own it is better to optimize it on problems that
polymorhpism cannot solve?

Page 14/Sec 5.2: I am a bit surprised about the way E is used and
the reasons it is used. Isn't Zipperposition an already fully
capable monomorphic superposition prover? Can you justify this?

"exclude Vampire because of parsing issues". Vampire should be fully
capable of parsing any compliant TPTP files and its parser is one
of the most efficient TPTP parsers.

----------------------- REVIEW 2 ---------------------
SUBMISSION: 17
TITLE: Iterative Monomorphisation
AUTHORS: Tanguy Bozec and Jasmin Blanchette

----------- Overall evaluation -----------
The paper describes an incremental monomorphisation algorithm, which can be
used to translate away rank-1 polymorphism from HOL/FOL problems by iteratively
instantiating the formulae’s type variables with concrete types. These problems
thus become accessible to reasoning systems that do not support polymorphism.
This necessarily incomplete but pragmatic approach has been briefly described
in Boehme's PhD thesis [5] and implemented as part of Isabelle/HOL’s SMT
integration and also used by the Sledgehammer system. The authors expand what
can be derived from these sources into a detailed description and suggest
guidelines for efficient implementation. They also implemented the algorithm in
the Zipperposition prover and report on its empirical evaluation on polymorphic
TPTP problems.

I am glad for a paper like this, as the topic clearly deserves a canonical
reference. At the same time, the question whether such format brings enough
contribution for a full conference paper cannot be avoided.

I was glad to see the exposition split into the "High level" and "Low level"
algorithm. The High level is already presented in a formally precise way and
accompanied by a small example, so what's left for the Low level are some
optimizations (make sure you don't repeat work of previous iterations; store
substitutions instead of formulas and only apply them in the end) and the
introduction of various numeric bounds that prevent the run from exploding (in
various dimensions). While establishing (maybe one could even say inventing)
these bounds seems important for practice, the optimizations, on the other
hand, look quite routine and their appearance in the pseudocode, which drives
the Low level section, make it (perhaps unnecessarily?) quite hard to penetrate
to a casual reader.

The evaluation part uses the new Zipperposition implementation of the
monomorphisation algorithm and describes how the authors decided on some
reasonable default values for the various numeric bounds to optimize
performance of the prover on the monomorphised versions of the polymorphic
problems from the TPTP library. Given the large number of these bounds, this
was a rather tedious process and the numbers reported in the numerous tables
are perhaps a bit uninteresting in the grand scheme of things. This again gives
the paper a bit of a workshop-ish flavor, but could arguably help future
implementors for the algorithm guide their way of tuning the bounds. I was also
not entirely convinced by the arguments for/against overfitting to the problem
set (details below).

Overall, however, given also how clearly the paper is written, I think it's
worth considering its acceptance!

Particulars:

- Could miniscoping help (or change the output to begin with)? It seems
  important what the individual formulas \varphi are in the input. What if we
  get \forall \alpha \in $tType. (\varphi_1 \land \varphi_2), could the
  algorithm benefit from instead considering? (\forall \alpha \in $tType.
  \varphi_1) \land (\forall \alpha \in $tType. \varphi_2)?

- "As a precaution against overfitting" - I don't see how this in itself
  could work as a precaution. Couldn't you still simply have overfitted to the
  500 random problems?

- Table 6 + the corresponding paragraph - so the "E timeout" is in seconds?
  And the "E call point" is a fraction of the overall timeout? "The longer
  Zipperposition runs before E is invoked, the more formulae are generated, and
  the more explosive iterative monomorphisation is." - how long does it run for
  a fraction value 0?

- "The results of these evaluations do not suggest that any overfitting took
  place during the option optimisation phase." - What's the justification for
  this? Could you please spell it out?

- "We unfortunately had to exclude Vampire because of parsing issues." - Did
  you try contacting the Vampire team about the issue or how to best mitigate
  it? They generally respond quite quickly these days.

----------------------- REVIEW 3 ---------------------
SUBMISSION: 17
TITLE: Iterative Monomorphisation
AUTHORS: Tanguy Bozec and Jasmin Blanchette

----------- Overall evaluation -----------
Paper summary
-------------

The paper introduces a method to generate monomorphic formulae out of
polymorphic formulae by substituting type variables by monomorphic types. This
is interesting in cases where an automatic theorem prover is to be invoked
which does not offer native support for polymorphism. In order to limit the
number of newly created formulae due to the instantiation of type variables,
several bounds are adopted. The approach is implemented in Zipperposition, a
polymorphic higher order prover. Zipperposition uses the E prover as a backend
and to date it could use E exclusively for monomorphic problems. This work
enables solving polymorphic problems in Zipperposition using E. An experimental
evaluation was carried out to determine the parameter combination on which
Zipperposition performed best on polymorphic formulae.

Overall evaluation
------------------

Applying the low level algorithm to Example 6 was hard. The notation is unclear
in parts and assumptions had to be made which might be incorrect (see also
questions to the authors), e.g.,:
- Function iterative_monomorphisation: S ← \emptyset, but S is never updated,
  instead S(\phi) is computed; similarly for N_old and N_old(\phi), N_new and
  N_new(\phi). I assume S is defined as the union of the S(\phi), same for
  N_old and N_new
- Function formula_mono_step: "(f |→ (υ_1,...,υ_n)) \in N_old(\phi) \union
  N_new(\phi)". I assume f is called on the non-monomorphic argument
  (υ_1,...,υ_n).
- Function formula_mono_step: M_next and N_next(\phi) are referenced, but they
  are defined in the main function iterative_monomorphisation and not known in
  formula_mono_step. I assume they are implicitly considered global variables.
- Function formula_mono_step: N_next and N_next(\phi). I assume N_next is the
  union of all N_next(\phi).
- p.7: "limit on the number of newly generated monomorphic type argument
  tupes". I assume this is max_mono_args in formula_mono_steps; "Similar bounds
  are used for the non-monomorphic type argument tuples". I assume this is
  max_nonm_args. However, their definitions do not coincide with the ones in
  formula_mono_steps.

The presentation need be more careful/consistent in some parts: 
- Function iterative_monomorphisation: the call of formula_mono_step does not
  match its definition (one parameter is missing, the parameter are ordered
  differently
- p.5: "match(v, \tau), where v and \tau are types"; from my understanding, v
  is a type variable
- Function formula_mono_step: type_variables is never used again (left hand
  side of "type_variables ← type_variables(\phi)"
- Function formula_mono_step: type_variables(\phi) vs.
  type_variables(v_1,...,v_n): argument type does not fit
- p.7: "limit on the number of newly generated non-/monomorphic type argument
  tupes". If these are max_mono_args/max_nonm_args, their definitions do not
  coincide with the ones in formula_mono_steps.
- p.7, second-to-last line: while loop -> foreach loop?

Adding more textual explanations and some examples would further enhance clarity.

The repository on Github seems not to be public, so I could not reproduce the
experiments.

I consider this work relevant. The presented approach leads to a significantly
better performance of Zipperposition with E compared to Zipperposition without
E. However, in my opinion the presentation need be enhanced significantly for
the paper to be accepted.

Questions to the authors
------------------------

- Function iterative_monomorphisation: how are S and S(\phi) related, and N_old
  and N_old(\phi), N_new and N_new(\phi)?
- p. 5: "match(v, \tau), where v and \tau are types"; isn't v a type variable
  instead of a type?
- Function formula_mono_step: what do mean bey "(f |→ (υ_1,...,υ_n))"?
- M_next and N_next(\phi): are these global variables?
- Which limits for generating argument tuples are you referring to on p. 7?
- p.7, 2nd par.: As far as I understood, no new formulae containing
  non-monomorphic type arguments are generated. However, your explanations
  suggest otherwise. What am I missing? Can you give an example?
- Function matches: wouldn't it be more efficient to check |S| before executing
  match(v_i, \tau_i) and to save the execution of match in case |S| is already
  to large?

Minor comments
--------------

- On a black and white printout, the blue text is barely distinguishable from
  the black text. I suggest to use a lighter shade.

------------------------------------------------------